\begin{problem}{2.3.1}
  For each of the following functions, $c=0$ lies on a periodic cycle. Classify
  this cycle as attracting, repelling, or neutral (non-hyperbolic). State if
  it is super attracting.
  \begin{align*}
    \text{i}.\  f(x) = \frac{\pi}{2} \cos(x),\quad \text{ii}.\ g(x) = -\frac{1}{2}x^3 - \frac{3}{2}x^2 + 1.
  \end{align*}
\end{problem}

\begin{proof}
  Recall that if $c$ is a point of period $r$, then $c$ is stable, asymptotically stable, unstable,
  if $f^r(c)$ is stable, asymptotically stable, unstable, respectively. Thus,
  if $c$ is a point of period $r$ and $f'(x)$ is continuous at $x=c$,
  then $c$ is asymptotically stable (attracting) if
  \begin{align*}
    \left|f'(f^0(c))\cdot f'(f^1(c)) \cdot \dots \cdot f'(f^{r-1}(c))\right| < 1
  \end{align*}
  and $c$ is unstable (repelling) if
  \begin{align*}
    \left|f'(f^0(c))\cdot f'(f^1(c)) \cdot \dots \cdot f'(f^{r-1}(c))\right| > 1.
  \end{align*}

  \begin{enumerate}
    \item Let $f(x) = \frac{\pi}{2} \cos(x)$. It is clear that $f^2(0) = 0$ so that $c=0$ is a period 2 point and
      $\{0, f(0)\}$ forms a 2-cycle.
      Note that $f'(x) = -\frac{\pi}{2}\sin(x)$, which is continuous, and that
      \begin{align*}
        \left|f'(0)\cdot f'(f(0))\right| &= \left|\left(-\frac{\pi}{2}\sin(0)\right)\left(-\frac{\pi}{2}\sin\left(\frac{\pi}{2}\right)\right)\right| = 0 < 1
      \end{align*}
      so that the 2-cycle $\{0, f(0)\}$ is asymptotically stable. Since
      \begin{align*}
        (f^2(0))' = (f(f(0)))' = f'(0)\cdot f'(f(0)) = 0,
      \end{align*}
      we have that $c=0$ is a super-attracting point of $f^2$ and the 2-cycle
      $\{0, f(0)\}$ is a super-attracting, asymptotically stable cycle.
    \item Let $g(x) = -\frac{1}{2}x^3 - \frac{3}{2}x^2 + 1.$ It is
      clear that $g^3(0) = 0$ so that $c=0$ is a period 3 point and $\{0, g(0), g^2(0)\}$
      forms a 3-cycle. Note that $g'(x) = -\frac{3}{2}x^2 - 3x$, which is continuous,
      and that
      \begin{align*}
        \left|g'(0)\cdot g'(g(0))\cdot g'(g^2(0))\right| &= \left|0 \left(-\frac{9}{2}\right)\left(\frac{3}{2}\right)\right| = 0 < 1
      \end{align*}
      so that the 2-cycle $\{0, g(0), g^2(0)\}$ is asymptotically stable. Since
      \begin{align*}
        (g^3(0))' = (g(g(g(0))))' = g'(0)\cdot g'(g(0))\cdot g'(g^2(0)) = 0,
      \end{align*}
      we have that $c=0$ is a super-attracting point of $g^3$ and the 3-cycle
      $\{0, g(0), g^2(0)\}$ is a super-attracting, asymptotically stable cycle.
  \end{enumerate}
\end{proof}
\newpage
