\begin{problem}{1.4.8}
  \begin{enumerate}
    \item Suppose that $f(c) = f'(c) = 0$ and $f''(c) \neq 0$. If $f''(x)$ is continuous at $x = c$,
      show that the Newton function $N_f (x)$ has a removable discontinuity at $x = c$. (Hint:
      Apply L'Hopital's rule to $N_f$ at $x = c$.)
    \item If in addition, $f'''(x)$ is continuous at $x = c$ with $f'''(c) \neq 0$,
      show that $N_f'(c) = 1/2$, so that $x = c$ is not a super-attracting fixed point in this case.
    \item Check the above for the function $f(x) = x^3 - x^2$ with $c = 0$.
  \end{enumerate}
\end{problem}

\begin{proof}
  \begin{enumerate}
    \item Suppose that $\lim_{x\to c} N_f(x) = L$. The function $N_f(x)$ will have a
      removable discontinuity at $x = c$ if $N_f(c)$ is undefined or $N_f(c) \neq L$ and
      the following function defined as
      \begin{align}\label{remove}
        F(x) :=
        \begin{cases}
          N_f(x) & \text{if $x \neq c$}\\
          L & \text{if $x = c$}
        \end{cases}
      \end{align}
      is continuous everywhere. Throughout, we assume without loss of generality that  $x=c$ is the only point such that $f'(x) = 0$.

      Note that if $x_0 \neq c$, then the differentiability of $f'$ implies that both $f$
      and $f'$ are continuous so that
      \begin{align*}
        \lim_{x\to x_0} N_f(x) = \lim_{x\to x_0} x - \frac{f(x)}{f'(x)} = x_0 - \frac{f(x_0)}{f'(x_0)} = N_f(x_0),
      \end{align*}
      i.e.\ the function $N_f(x)$ is continuous.

      On the other hand, at the point $x = c$, we see from the continuity of $f$ and $f'$
      that
      \begin{align*}
        \lim_{x\to c} N_f(x) &= \lim_{x\to c} x - \frac{f(x)}{f'(x)} \\
        &= \lim_{x\to c} \frac{xf'(x) - f(x)}{f'(x)} \\
        &= \frac{c f'(c) - f(c)}{f'(c)}.
      \end{align*}
      From our supposition that $f(c) = f'(c) = 0$, we see that $c f'(c) - f(c) = 0$ and $f'(c) = 0$
      implying that $\lim_{x\to c} N_f(x) = 0/0$, an indeterminate form. By L'Hospital's rule
      in conjunction with our supposition that $f''$ is continuous and $f''(c) \neq 0$,
      since $\lim_{x\to c} x f'(x) - f(x) = \lim_{x\to c} f'(x) = 0$, we know that
      \begin{align*}
        \lim_{x\to c} N_f(x) &= \lim_{x\to c} \frac{xf'(x) - f(x)}{f'(x)} \\
        &= \lim_{x\to c} \frac{[xf'(x) - f(x)]'}{[f'(x)]'} \\
        &= \lim_{x\to c} \frac{x f''(x)}{f''(x)} = c.
      \end{align*}

      Therefore, the function
      \begin{align*}
        F(x) :=
        \begin{cases}
          N_f(x) & \text{if $x \neq c$}\\
          c & \text{if $x = c$}
        \end{cases}
      \end{align*}
      is continuous everywhere and $x = c$ is a removable discontinuity for the
      function $N_f(x)$.

    \item Suppose that $f'''(x)$ is continuous at $x = c$ with $f'''(c) \neq 0$. From the definition of $N_f(x)$,
      we see that
      \begin{align*}
        N_f'(x) = \left[\frac{xf'(x) - f(x)}{f(x)}\right]' = \frac{f(x)f''(x)}{f'(x)^2}.
      \end{align*}
      Thus, we have from our initial assumptions that
      \begin{align*}
        \lim_{x\to c} N_f'(x) = \lim_{x\to c}\frac{f(x)f''(x)}{f'(x)^2} = \frac{0}{0},
      \end{align*}
      an indeterminate form. The assumptions to use L'Hospital's rule are met and we see
      from the continuity of $f$ and its derivatives combined with our initial suppositions that
      \begin{align*}
        \lim_{x\to c} N_f'(x) &= \lim_{x\to c}\frac{[f(x)f''(x)]'}{[f'(x)^2]'} \\
        &= \lim_{x\to c}\frac{f'(x)f''(x) + f(x)f'''(x)}{2f'(x)f''(x)} = \frac{0}{0},
      \end{align*}
      another indeterminate form. Applying L'Hospital's rule one last time we see
      that
      \begin{align*}
        \lim_{x\to c} N_f'(x) &= \lim_{x\to c}\frac{[f'(x)f''(x) + f(x)f'''(x)]'}{[2f'(x)f''(x)]'}\\
        &= \lim_{x\to c}\frac{f''(x)^2 + 2f'(x)f'''(x) + f(x) f^{(4)}(x)}{2f''(x)^2 + 2 f'(x)f'''(x)} = \frac{1}{2}.
      \end{align*}
      Therefore, since $|N_f'(c)| \neq 0$, the point $x=c$ is not a super-attracting
      fixed point.
    \item Let $f(x) = x^3 - x^2$. Note that $f'(x) = 3x^2 - 2x$, $f''(x) = 6x - 2$, and $f'''(x) = 6$, all of which are continuous.
      Further, we know that
      \begin{align*}
        N_f(x) = x - \frac{f(x)}{f'(x)} = \frac{2x^2 - x}{3x - 2}.
      \end{align*}
      It is clear that the point $c = 0$ is a fixed point of the function $f$. Note that
      $f(c) = f'(c) = 0$ with $f''(c) \neq 0$ and $f'''(c) \neq 0$. However, the point $x=c$ is
      not a removable discontinuity since $\lim_{x\to c}N_f(x) = 0 = N_f(c)$, i.e.\
      the Newton function is actually continuous at $x=c$ so there is no discontinuity to remove. Since $|N_f'(c)| = 1/2$,
      the result that $c = 0$ is not a super-attracting fixed point still holds.

  \end{enumerate}
\end{proof}
\newpage
