\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amsthm, amssymb, amsfonts, enumitem, graphicx}
\usepackage{fancyhdr}

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\renewcommand*{\proofname}{Solution}
\renewcommand{\theenumi}{\alph{enumi}}

\newcommand{\vect}[1]{\boldsymbol{#1}}

\newenvironment{amatrix}[1]{%
  \left[\begin{array}{@{}*{#1}{c}|c@{}}
}{%
  \end{array}\right]
}

\newcommand{\grstep}[2][\relax]{%
   \ensuremath{\mathrel{
       {\mathop{\longrightarrow}\limits^{#2\mathstrut}_{
                                     \begin{subarray}{l} #1 \end{subarray}}}}}}
\newcommand{\swap}{\leftrightarrow}

\pagestyle{fancy}
\fancyhf{}
\rhead{Homework Assignment 2}
\lhead{Matthew Tiger}
\cfoot{\thepage}


\title{Homework Assignment 2}
\author{Matthew Tiger}


\begin{document}


\maketitle


% Problem 1
\begin{problem}
  Convert the following linear programming problem to \textit{standard form}:
  \begin{align*}
    \begin{array}{rll}
      \text{maximize} & 2x_1 + x_2 &\\
      \text{subject to} & 0 \leq x_1 &\leq 2 \\
      & x_1 + x_2 &\leq 3 \\
      & x_1 + 2x_2 &\leq 5 \\
      & x_2 \geq 0 &
    \end{array}
  \end{align*}
\end{problem}

\begin{proof}
  In order to convert this linear programming problem into standard form, we
  must transform the objective from \textit{maximize} to \textit{minimize} and
  the constraints must be transformed from linear inequalities into linear equations.

  Our first step will be to rewrite the objective function as a minimization problem
  and write each constraint as a linear inequality as so:
  \begin{align*}
    \begin{array}{rll}
      \text{minimize} & -2x_1 - x_2 &\\
      \text{subject to} & x_1 &\leq 2 \\
      & x_1 + x_2 &\leq 3 \\
      & x_1 + 2x_2 &\leq 5 \\
      & x_1 \geq 0, x_2 \geq 0 &
    \end{array}
  \end{align*}

  We can then introduce three slack variables $x_3, x_4, x_5$ to turn the linear
  inequalities into linear equations:
  \begin{align*}
    \begin{array}{rllllll}
      \text{minimize} & -2x_1 &- x_2  & & & &\\
      % \text{subject to} & x_1 & &+ x_3 & & &= 2 \\
      \text{subject to} & x_1 &+ x_2 &+ x_3 & & &= 2 \\
      & x_1 &+ x_2 & &+ x_4 & &= 3 \\
      & x_1 &+ 2x_2 & & &+ x_5&= 5 \\
      & x_1 \geq 0, &x_2 \geq 0, &x_3 \geq 0, &x_4 \geq 0, &x_5 \geq 0
    \end{array}
  \end{align*}

  As the above linear programming problem is written as
  \begin{align*}
    \begin{array}{rl}
      \text{minimize} & \vect{c}^\intercal \vect{x}\\
      \text{subject to} & A\vect{x} = \vect{b} \\
      & \vect{x} \geq 0
    \end{array}
  \end{align*}
  where
  \begin{align*}
    \vect{c}^\intercal =
    \begin{bmatrix}
      -2 \\
      -1 \\
    \end{bmatrix}^\intercal,
    \quad
    A =
    \begin{bmatrix}
      1 & 1 & 1 & 0 & 0 \\
      1 & 1 & 0 & 1 & 0 \\
      1 & 2 & 0 & 0 & 1 \\
    \end{bmatrix},\quad
    \vect{x} =
    \begin{bmatrix}
      x_1 \\
      x_2 \\
      x_3 \\
      x_4 \\
      x_5 \\
    \end{bmatrix}, \quad
    \vect{b} =
    \begin{bmatrix}
      2 \\
      3 \\
      5 \\
    \end{bmatrix}
  \end{align*}
  with $\vect{x} \geq 0$ and $\vect{b} \geq 0$ the linear programming problem
  is in standard form and we are done.
\end{proof}
\newpage


% Problem 2
\begin{problem}
  Solve the system $A\vect{x} = \vect{b}$ where
  \begin{align*}
    A =
    \begin{bmatrix}
      2 & -1 & 2 & -1 & 3 \\
      1 & 2 & 3 & 1 & 0 \\
      1 & 0 & -2 & 0 & -5 \\
    \end{bmatrix}
    , \quad
    \vect{b} = \begin{bmatrix}
      14 \\
      5 \\
      -10 \\
    \end{bmatrix}.
  \end{align*}
  If possible, generate a non-basic feasible solution of the system from which
  you derive next a basic feasible one.
\end{problem}

\begin{proof}
  In order to solve the system $Ax=b$, we must perform row operations on the
  augmented matrix to reduce it to reduced row form. We perform these operations below:

  \begin{align*}
    &\begin{amatrix}{5}
      2 & -1 & 2 & -1 & 3 & 14 \\
      1 & 2 & 3 & 1 & 0 & 5 \\
      1 & 0 & -2 & 0 & -5 & -10 \\
    \end{amatrix}
    \begin{array}{c}
    \grstep[{[2] - (1/2)[1]}]{(1/2)[1]} \\
    \grstep[]{[3] - (1/2)[1]}
    \end{array}
    \begin{amatrix}{5}
      1 & -1/2 & 1 & -1/2 & 3/2 & 7 \\
      0 & 5/2 & 2 & 3/2 & -3/2 & -2 \\
      0 & 1/2 & -3 & 1/2 & -13/2 & -17 \\
    \end{amatrix}
    \begin{array}{c}
      \grstep[]{(2/5)[2]}
    \end{array}
    \\
    &\begin{amatrix}{5}
      1 & -1/2 & 1 & -1/2 & 3/2 & 7 \\
      0 & 1 & 4/5 & 3/5 & -3/5 & -4/5 \\
      0 & 1/2 & -3 & 1/2 & -13/2 & -17 \\
    \end{amatrix}
    \begin{array}{c}
      \grstep[{[3] - (1/2)[2]}]{[1] + (1/2)[2]} \\
    \end{array}
    \begin{amatrix}{5}
      1 & 0 & 7/5 & -1/5 & 6/5 & 33/5 \\
      0 & 1 & 4/5 & 3/5 & -3/5 & -4/5 \\
      0 & 0 & -17/5 & 1/5 & -31/5 & -83/5 \\
    \end{amatrix}
    \\
    &\begin{array}{c}
      \grstep[]{(-5/17)[3]}
    \end{array}
    \begin{amatrix}{5}
      1 & 0 & 7/5 & -1/5 & 6/5 & 33/5 \\
      0 & 1 & 4/5 & 3/5 & -3/5 & -4/5 \\
      0 & 0 & 1 & -1/17 & 31/17 & 83/17 \\
    \end{amatrix}
    \begin{array}{c}
      \grstep[{[2] - (4/5)[3]}]{[1] - (7/5)[3]} \\
    \end{array}
    \begin{amatrix}{5}
      1 & 0 & 0 & -2/17 & -23/17 & -4/17 \\
      0 & 1 & 0 & 11/17 & -35/17 & -80/17 \\
      0 & 0 & 1 & -1/17 & 31/17 & 83/17 \\
    \end{amatrix}.
  \end{align*}

  Using the above row-reduced augmented matrix, we see that the solution to the
  system $A\vect{x} = \vect{b}$ is given by
  \begin{align}\label{prob_2_x}
    \vect{x} =
    \begin{bmatrix}
      -4/17 \\
      -80/17 \\
      83/17 \\
      0 \\
      0 \\
    \end{bmatrix}
    +
    s
    \begin{bmatrix}
      2/17 \\
      -11/17 \\
      1/17 \\
      1 \\
      0 \\
    \end{bmatrix}
    +
    t
    \begin{bmatrix}
      23/17 \\
      35/17 \\
      -31/17 \\
      0 \\
      1 \\
    \end{bmatrix}
  \end{align}
  where $s\in\mathbb{R}$ and $t\in\mathbb{R}$.

  Suppose that the matrix $A$ is written such that $A = [\vect{a}_i]$ for
  $1 \leq i \leq 5$ where $\vect{a}_i$ corresponds to the $i$-th column of the original
  matrix $A$. Recall that a solution $\vect{x}_0 \geq \vect{0}$ of the system $A\vect{x} = \vect{b}$ is a
  basic feasible solution if the columns of the matrix
  $A$ associated to the nonzero components of $\vect{x}_0$ are linearly
  independent. Otherwise the solution is a non-basic feasible solution.

  Using solution \eqref{prob_2_x}, we see that for $s = 0$ and $t = 82/31$, we get
  the corresponding feasible solution $\vect{x}_0 = [1762/527, 390/527, 1/17, 0, 82/31]^\intercal$ to the system
  $A\vect{x} = \vect{b}$. We know that this is a non-basic feasible solution since the vectors
  $\vect{a}_1$, $\vect{a}_2$, $\vect{a}_3$, and $\vect{a}_5$ must be linearly
  dependent as the $\text{rank}(A) = 3$, i.e.\ the maximum number of linearly independent columns of $A$ is 3.

  The Fundamental Theorem of LP prescribes how to move from this non-basic feasible
  solution $\vect{x}_0$ to a basic feasible solution $\vect{x}_1$. As $\vect{a}_1$, $\vect{a}_2$, $\vect{a}_3$, and $\vect{a}_5$
  are linearly dependent, there exists constants $y_1, y_2, y_3, y_5$ not all zero such that
  \begin{align*}
    y_1\vect{a_1} + y_2\vect{a_2} + y_3\vect{a_3} + y_5\vect{a_5} = \vect{0},
  \end{align*}
  namely $y_1 = 1$, $y_2 = 35/23$, $y_3 = -31/23$, and $y_5 = 17/23$.
  Thus, the vector $\epsilon\vect{y} = \epsilon[y_1, y_2, y_3, 0, y_5] ^ \intercal$ satisfies
  $A[\epsilon\vect{y}] = \vect{0}$. As such, the vector $\vect{x}_0 - \epsilon\vect{y}$ satisfies
  $A[\vect{x}_0 - \epsilon\vect{y}] = \vect{b}$, i.e.\ the vector $\vect{x}_0 - \epsilon\vect{y}$
  is a solution of the original system $A\vect{x} = \vect{b}$. Choose
  $$\epsilon = \min\{x_i/y_i| i=1,2,3,5\ y_i > 0\} = -23/527.$$
  Then the vector $\vect{x}_1 = \vect{x}_0 - \epsilon\vect{y}$ will have 3 positive components and the rest of the components will be 0 showing that
  the vector $\vect{x}_1$ is a basic feasible solution. Therefore,
  \begin{align*}
    \vect{x}_1 = \vect{x}_0 - \epsilon\vect{y} = \begin{bmatrix}105/31\\ 25/31\\ 0\\ 0\\ 83/31\end{bmatrix}
  \end{align*}
  is the desired basic feasible solution.
\end{proof}
\newpage


% Problem 3
\begin{problem}
  Does every linear programming problem in standard form have a nonempty feasible set?
  If ``yes'', provide a proof. If ``no'', provide a counter-example.

  Does every linear programming problem in standard form (assuming a nonempty feasible
  set) have an optimal solution? If ``yes'', provide a proof. If ``no'', provide a counter-example.
\end{problem}

\begin{proof}
\end{proof}
\newpage


% Problem 4
\begin{problem}
  \begin{enumerate}
    \item Solve the following linear program graphically:
      \begin{align*}
        \begin{array}{ll}
          \text{maximize} & 2x_1 + 5x_2 \\
          \text{subject to} & 0 \leq x_1 \leq 4 \\
          & 0 \leq x_2 \leq 6 \\
          & x_1 + x_2 \leq 8
        \end{array}
      \end{align*}
    \item Solve the linear program in (b) the same way Example 15.15 was solved in class.
      Compute only the vertices that lead to the optimal vertex found at (a).
  \end{enumerate}
\end{problem}

\begin{proof}
\end{proof}

\end{document}
