\begin{problem}{3}
  Consider the family of functions $f_\lambda(x) = x^3 - \lambda x$ for some parameter
  $\lambda \in \mathbb{R}$.

  \begin{enumerate}
    \item Find all fixed points and determine their nature and where they
      are created as $\lambda$ varies.
    \item Find where a 2-cycle is created
      and give the graph of where this happens. Determine the stability of the
      hyperbolic 2-cycles.
    \item Use a graphing utility to find an approximate
      value of $\lambda$ where the 3-cycle is created. Give the graph of this situation.
  \end{enumerate}
\end{problem}

\begin{proof}
  \begin{enumerate}
    \item The fixed points of $f_\lambda$ are the roots of the function
      \begin{align*}
        g_\lambda(x) &= f_\lambda(x) - x\\
        &= x(x^2 - \lambda  - 1).
      \end{align*}
      Thus, the fixed points of $f_\lambda$ are $x_0 = 0$, $x_1 = \sqrt{\lambda + 1}$, and $x_2 = -\sqrt{\lambda + 1}$.
      Note that the points $x_1$ and $x_2$ are real only if $\lambda \geq -1$, i.e.\
      the points are only fixed points of the dynamical system if $\lambda \geq -1$.

      Using the first derivative of $f_\lambda$, we can classify the above fixed points
      when they are hyperbolic. If the fixed point is non-hyperbolic, we can use
      the second and third derivatives when the fixed point is non-hyperbolic
      of the type $f_\lambda'(x) = 1$, and the Schwarzian derivative when the fixed point is non-hyperbolic
      of the type $f_\lambda'(x) = -1$.      Note that
      \begin{align*}
        f_\lambda'(x) &= 3x^2 - \lambda \\
        f_\lambda''(x) &= 6x \\
        f_\lambda'''(x) &= 6.
      \end{align*}
      If $f_\lambda'(x) = -1$, we see that the Schwarzian derivative of $f_\lambda$ is given by
      \begin{align*}
        Sf_\lambda(x) &= -f_\lambda'''(x) - \frac{3}{2} \left[f_\lambda''(x)\right]^2 \\
        &= -6 - 54x^2.
      \end{align*}

      For the fixed point $x_0 = 0$, we see that $|f_\lambda'(x_0)| = |\lambda|$. Thus, the fixed point $x_0$ is
      a hyperbolic fixed point if $\lambda \neq -1$ or $\lambda \neq 1$. If $|\lambda| < 1$, then $x_0$ is asymptotically stable
      and if $|\lambda| > 1$, then $x_0$ is an unstable fixed point. If $\lambda = -1$, then $f_\lambda'(x_0) = 1$.
      Since $f_\lambda''(x_0) = 0$ and $f_\lambda'''(x_0) = 6 > 0$, the fixed point $x_0$ is unstable. If $\lambda = 1$,
      then $f_\lambda'(x_0) = -1$. The Schwarzian derivative of $f_\lambda$ at $x_0$ is then $Sf_\lambda(x_0) = -6 < 0$. Therefore,
      the fixed point $x_0$ is an asymptotically stable fixed point.

      Consider now the fixed point $x_1 = \sqrt{\lambda + 1}$ for $\lambda \geq -1$. We readily see that $|f_\lambda'(x_1)| = |3 + 2\lambda|$.
      If $\lambda > -1$, then $|f_\lambda'(x_1)| > 1$ and $x_1$ is
      hyperbolic and unstable. If $\lambda = -1$, then $x_1 = 0 = x_0$ and from the previous
      classification of the fixed point $x_0$, we know that $x_1$ is unstable.

      Lastly, consider the fixed point $x_2 = -\sqrt{\lambda + 1}$ for $\lambda \geq -1$.
      We thus have that $|f_\lambda'(x_2)| = |3 + 2\lambda|$ and the same classification for $x_1$ holds for $x_2$, i.e.\
      the fixed point $x_2$ is hyperbolic and unstable if $\lambda > -1$ and non-hyperbolic and unstable if $\lambda = -1$.

    \item Recall that a point $x$ is a period 2 point of $f_\lambda$ if $f_\lambda^2(x) = x$ and $f_\lambda(x) \neq x$. The 2-cycle associated to the period 2 point
      is then $\{x, f_\lambda(x)\}$. We thus look for solutions to the equation
      \begin{align}\label{period2}
        f_\lambda^2(x) - x &= \left(x^3 - \lambda x\right)^3 - \lambda \left(x^3 - \lambda x\right) - x \notag\\
        &= x^9 -3 \lambda x^7 + 3 \lambda ^2 x ^5 - \lambda ^3 x^3 - \lambda x^3 + \lambda^2 x - x \notag\\
        &= x (x^4- \lambda x^2 + 1)(x^2 - \lambda -1) (x^2 - \lambda + 1) = 0.
      \end{align}
      Suppose first that $\lambda < -1$. Then the only fixed point of the function $f_\lambda$
      is $x_0 = 0$ so that $x=0$ can be factored out of \eqref{period2} since the solutions we seek satisfy $f_\lambda(x) \neq x$. After factoring $x$ out from the above polynomial we have that
      \begin{align*}
        (x^4- \lambda x^2 + 1)(x^2 - \lambda -1) (x^2 - \lambda + 1) = 0.
      \end{align*}
      However, if $\lambda < -1$, then $(x^4- \lambda x^2 + 1) = 0$, $(x^2 - \lambda -1) = 0$,
      and $(x^2 - \lambda + 1) = 0$, all have no real solutions. Therefore, if $\lambda < -1$,
      then $f_\lambda$ has no period 2 points.

      Now consider $\lambda \geq -1$. Then for similar reasons we can factor $(x - x_0)(x - x_1)(x - x_2)$
      out of \eqref{period2} and thus see that
      \begin{align*}
        (x^4- \lambda x^2 + 1)(x^2 - \lambda + 1) = 0
      \end{align*}

      To continue, first polynomial has real solutions only if $\lambda \geq 2$. Second polynomial only has real solutions if $\lambda \geq 1$.
      Thus $-1 \leq \lambda \leq 1$ there are no period 2 points.

    \item
  \end{enumerate}
\end{proof}
\newpage
