\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amsthm, amssymb, amsfonts, enumitem, graphicx}
\usepackage{fancyhdr}
\usepackage{breqn}

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\renewcommand*{\proofname}{Solution}
\renewcommand{\theenumi}{\alph{enumi}}

\newcommand{\vect}[1]{\boldsymbol{#1}}

\newenvironment{case}{%
  \let\olditem\item%
  \renewcommand\item[1][]{\olditem \textbf{##1} \\}%
  \begin{enumerate}[label=\textbf{Case \arabic*:},itemindent=*,leftmargin=0em]}{\end{enumerate}%
}

\pagestyle{fancy}
\fancyhf{}
\rhead{Homework Assignment 4}
\lhead{Matthew Tiger}
\cfoot{\thepage}


\title{Homework Assignment 4}
\author{Matthew Tiger}


\begin{document}


\maketitle


% Problem 1
\begin{problem}
  Find the first three terms in the asymptotic expansions of $x \to 0^+$ of the following
  integrals:
  \begin{align*}
    \int_x^1 \cos(xt)dt, \qquad \int_0^{1/x} e^{-t^2}dt.
  \end{align*}
\end{problem}

\begin{proof}
\end{proof}
\newpage


% Problem 2
\begin{problem}
  Find the full asymptotic behavior as $x \to 0^+$ of the following integral:
  \begin{align*}
    \int_{0}^1 \frac{e^{-t}}{1 + x^2t^3}dt
  \end{align*}
\end{problem}

\begin{proof}
\end{proof}
\newpage

% Problem 3
\begin{problem}
  Find the full asymptotic expansion of $\int_0^x \text{Bi}(t) dt$ as $x \to +\infty$.
\end{problem}

\begin{proof}
\end{proof}
\newpage


% Problem 4
\begin{problem}
  Find the first five terms in the asymptotic expansion as $x \to +\infty$ of the integral
  \begin{align*}
    \int_0^{\pi/4} e^{-xt^2}\sqrt{\tan t}   dt
  \end{align*}
  \begin{enumerate}
    \item by using a suitable change of variables and then applying Watson's lemma.
    \item by applying Laplace's method directly to the given integral.
  \end{enumerate}
\end{problem}

\begin{proof}
\end{proof}
\newpage


% Problem 5
\begin{problem}
  Use Laplace's method of moving maxima to obtain the first two terms in the asymptotic
  expansion as $x \to +\infty$ of the integral
  \begin{align}\label{exp_int}
    \int_{0}^\infty \exp\left[-t - \frac{x}{\sqrt{t}}\right] dt.
  \end{align}
\end{problem}

\begin{proof}
\end{proof}
\newpage


% Problem 6
\begin{problem}
  Let $f(x, t)$ be differentiable in $x$ and continuous in $(x, t)$ on $I \times J$,
  where $I$ and $J$ are intervals, and suppose that there exist functions $g(t)$ and $g_1(t)$
  that are integrable on $J$ such that for all $(x,t) \in I \times J$ we have that
  \begin{align*}
    |f(x,t)|\leq g(t) \quad \text{and} \quad |\partial_x f(x,t)| \leq g_1(t).
  \end{align*}
  Then
  \begin{align*}
    \frac{d}{dx} \int_J f(x,t) dt = \int_J \partial_x f(x,t) dt.
  \end{align*}
  \begin{enumerate}
    \item Let $0 < a < b < \infty$. Use the above theorem to show that if $x \in (a,b)$, then
      \begin{align*}
        \frac{d^3}{dx^3} \int_0^\infty \exp\left[-t - \frac{x}{\sqrt{t}}\right] dt = -\int_0^\infty t^{-3/2}\exp\left[-t - \frac{x}{\sqrt{t}}\right] dt.
      \end{align*}
    \item Use integration by parts to show that
      \begin{align*}
        \int_0^\infty \exp\left[-t - \frac{x}{\sqrt{t}}\right] dt = \frac{x}{2}\int_0^\infty t^{-3/2}\exp\left[-t - \frac{x}{\sqrt{t}}\right] dt.
      \end{align*}
    \item Combine parts (a) and (b) to prove that integral \eqref{exp_int} is a solution of the
      differential equation $xy''' + 2y = 0$ that also satisfies the initial condition $y(0) = 1$.
      Then use integration by parts to give an easy direct proof that the integral also satisfies
      the condition $y(+\infty) = 0$.
  \end{enumerate}
\end{problem}

\begin{proof}
\end{proof}
\newpage


% Problem 7
\begin{problem}
  \begin{enumerate}
    \item Find the leading behavior as $x\to + \infty$ of Laplace integrals of the form
      \begin{align*}
        \int_a^b (t-a) ^\alpha g(t) e^{x\phi(t)} dt
      \end{align*}
      where $\phi(t)$ has a maximum at $t = a$, $g(a) = 1$. Suppose further that $\alpha > -1$ and $\phi'(a) < 0$.
    \item Repeat the analysis of part (a) when $\alpha > -1$
      and $\phi'(a) = \phi''(a) = \dots = \phi^{(p-1)}(a) = 0$ and $\phi^{(p)}(a) < 0$.
  \end{enumerate}
\end{problem}

\begin{proof}
\end{proof}
\newpage


\end{document}